---
title: "Supplementary tables 3-11"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
author: "Liang-Yu Lin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r include=FALSE}
# Load data and packages
# Datasets

# packages
library(markdown)
library(rmarkdown)
library(arsenal)
library(kableExtra)
library(finalfit) 
library(survival)
library(survminer)

# Limit significant digits to 2, remove scientific notation
options(digits = 2, scipen = 999)
# Time cut points
time_cut <- c(224) # British summer time 
```


# Vitamin D prescription

Exposure: had vitamin D prescription records (variable name: vd_drug)

## Supplementation table 3: vd_drug and covid19 diagnosis

Use s`urvsplit` to split the follow-up time
```{r echo=FALSE}
bd_d_s <- survSplit(
  Surv(bd_d$fu_time, event = bd_d$c_infection)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_d",
  start = "start",
  episode = "season"
)
```


### Complete regression table
```{r echo=FALSE}
explanatory = c("vd_drug*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")
explanatory_multi = c("vd_drug*strata(season)", "sex", "age_group")
dependent ="Surv(start, end_d, event = event)"

bd_d_s %>% 
  finalfit(dependent, explanatory, explanatory_multi, keep_models = T) %>% 
knitr::kable(format = "pandoc")

```

### Multiple regression: 
reason: the final fit output is messy in displaying more than two variables with interaction. 
```{r include=FALSE}
coxph(Surv(start, end_d, event = event)~vd_drug*strata(season)+sex+age_group+ethnic_c+bmi_group+drink_freq_c + smoke_stat+ bst+ imd_bd_q+ regions+ low_immune+ chronic_disease, 
      data = bd_d_s, id = f.eid) %>% summary()
```

### Simple regression:
```{r echo=FALSE}
coxph(Surv(start, end_d, event = event)~vd_drug*strata(season), data = bd_d_s, id = f.eid) %>% summary()
```

### Partial adjusted for sex and age group
```{r echo=FALSE}
# write a function
partial_sen_o1 <- function(x){
  bd_d_s %>% 
  finalfit(dependent ="Surv(start, end_d, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune",  "chronic_disease")

map(var_list, partial_sen_o1)

```


## Supplementation table 4: vd_drug and covid-19 admission

```{r include=FALSE}
# data management

bd_d$fu_time_t3 <- (as.numeric(bd_d$end_hos_all)-as.numeric(start))

# Split the following time:

bd_d_t3 <- survSplit(
  Surv(bd_d$fu_time_t3, event = bd_d$hos_all)~.,
  data = bd_d,
  cut = time_cut, 
  end = "end_hos_all",
  start = "start",
  episode = "season"
)
```

### Complete table
```{r echo=FALSE}
explanatory = c("vd_drug*strata(season)","sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat" ,"bst", "imd_bd_q" , "regions","low_immune", "chronic_disease")
explanatory_multi = c("vd_drug*strata(season)", "sex", "age_group")
# survival objects: 
# time-in:  2020-03-15
# time-out: end_hos

dependent ="Surv(start, end_hos_all, event = event)"

# model
bd_d_t3 %>% finalfit(dependent, explanatory, explanatory_multi, keep_models = T) %>% knitr::kable(format = "pandoc", row.names = NA)
```

### multiple regression:
```{r echo=FALSE}
coxph(
  Surv(start, end_hos_all, event =event) ~ vd_drug*strata(season)+sex+age_group+ethnic_c+bmi_group+drink_freq_c+smoke_stat+bst+imd_bd_q+regions+low_immune+chronic_disease, 
  data = bd_d_t3 , id = f.eid) %>% summary()
```

### Simple regression:
```{r echo=FALSE}
coxph(Surv(start, end_hos_all, 
           event =event) ~ vd_drug*strata(season),
      data = bd_d_t3 , id = f.eid) %>% summary()
```

### Partial adjustment
```{r echo=FALSE}
# function
partial_sen_o2 <- function(x){
  bd_d_t3 %>% 
  finalfit(dependent ="Surv(start, end_hos_all, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_o2)

```


## Supplementation table 5: vd_drug and mortality

Data management:
```{r include=FALSE}
bd_d$fu_time_t4 <-(as.numeric(bd_d$end_mor_all) - as.numeric(start))
# Split the following time:

bd_d_t4 <- survSplit(
  Surv(bd_d$fu_time_t4, event = bd_d$n_death_covid_all)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_mor_all",
  start = "start",
  episode = "season"
)

```


### Full table
```{r echo=FALSE}
explanatory = c("vd_drug*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst" ,"imd_bd_q", "regions", "low_immune", "chronic_disease")
explanatory_multi = c("vd_drug*strata(season)", "sex", "age_group")

dependent ="Surv(start, end_mor_all, event = event)"

# model
bd_d_t4 %>%  finalfit(dependent,explanatory) %>% knitr::kable(format = "pandoc", row.names = NA)
```

### Multiple regression
```{r eval=FALSE, include=FALSE}
coxph(Surv(start, end_mor_all, event = event)~ vd_drug*strata(season)+sex+age_group+ethnic_c+bmi_group+drink_freq_c+smoke_stat+bst+imd_bd_q+regions+low_immune+chronic_disease, 
      data = bd_d_t4 , id = f.eid) %>% summary()
```

### Simple regression
```{r echo=FALSE}
coxph(Surv(start, end_mor_all, event = event)~vd_drug*strata(season), data = bd_d_t4 , id = f.eid) %>% summary()
```


### Partial adjusted for sex and age
for vitamin D
```{r include=FALSE}
coxph(Surv(start, end_mor_all, event = event)~ vd_drug*strata(season)+sex+age_group, 
      data = bd_d_t4 , id = f.eid) %>% summary()
```

```{r include=FALSE}
partial_sen_o3 <- function(x){
  bd_d_t4 %>% 
  finalfit(dependent ="Surv(start, end_mor_all, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_o3)
```



# Self-report vitamin D supplementation

## Supplementation table 6: vitd_supp and covid19 infection

Use `survsplit` to split the follow-up time
```{r include=FALSE}
bd_d_s <- survSplit(
  Surv(bd_d$fu_time, event = bd_d$c_infection)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_d",
  start = "start",
  episode = "season"
)
```

### Complete regression table
```{r echo=FALSE}
explanatory = c("vitd_supp*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")
explanatory_multi = c("vitd_supp*strata(season)", "sex", "age_group")
dependent ="Surv(start, end_d, event = event)"

bd_d_s %>% 
  finalfit(dependent, explanatory, explanatory_multi, keep_models = T) %>% 
knitr::kable(format = "pandoc")

```

### Multiple regression
```{r echo=FALSE}
coxph(Surv(start, end_d, event = event) ~         vitd_supp*strata(season)+sex+age_group+ethnic_c+bmi_group+drink_freq_c+smoke_stat+bst+imd_bd_q+regions+low_immune+chronic_disease, 
      data = bd_d_s, id = f.eid) %>% summary()
```

### Simple regression:
```{r echo=FALSE}
coxph(Surv(start, end_d, event = event)~vitd_supp*strata(season), data = bd_d_s, id = f.eid) %>% summary()
```

### Partial adjusted for sex and age group
```{r echo=FALSE}
# write a function
partial_sen_o1 <- function(x){
  bd_d_s %>% 
  finalfit(dependent ="Surv(start, end_d, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("vitd_supp*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_o1)

```

## Supplementation table 7: vitd_supp and covid-19 admission

```{r include=FALSE}
# data management

bd_d$fu_time_t3 <- (as.numeric(bd_d$end_hos_all)-as.numeric(start))

# Split the following time:

bd_d_t3 <- survSplit(
  Surv(bd_d$fu_time_t3, event = bd_d$hos_all)~.,
  data = bd_d,
  cut = time_cut, 
  end = "end_hos_all",
  start = "start",
  episode = "season"
)
```

### Full table
```{r echo=FALSE}
explanatory = c("vitd_supp*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst", "imd_bd_q" , "regions","low_immune", "chronic_disease")
explanatory_multi = c("vitd_supp*strata(season)", "sex", "age_group")
# survival objects: 
# time-in:  2020-03-15
# time-out: end_hos

dependent ="Surv(start, end_hos_all, event = event)"

# model
bd_d_t3 %>% finalfit(dependent, explanatory, explanatory_multi, keep_models = T) %>% knitr::kable(format = "pandoc", row.names = NA)
```

### Multiple regression
```{r eval=FALSE, include=FALSE}
coxph(Surv(start, end_hos_all, event = event) ~ vitd_supp*strata(season) + sex + age_group + ethnic_c + bmi_group + drink_freq_c + smoke_stat + bst+imd_bd_q + regions + low_immune + chronic_disease, 
      data = bd_d_t3 , id = f.eid) %>% summary()
```

### Simple regression:
```{r echo=FALSE}
coxph(Surv(start, end_hos_all, event = event)~vitd_supp*strata(season), data = bd_d_t3 , id = f.eid) %>% summary()
```

### Partial adjustment
```{r echo=FALSE}
# function
partial_sen_o2 <- function(x){
  bd_d_t3 %>% 
  finalfit(dependent ="Surv(start, end_hos_all, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_o2)

```

## Supplementation table 8: vitd_supp and mortality

Data management:
```{r include=FALSE}
bd_d$fu_time_t4 <-(as.numeric(bd_d$end_mor_all) - as.numeric(start))
# Split the following time:

bd_d_t4 <- survSplit(
  Surv(bd_d$fu_time_t4, event = bd_d$n_death_covid_all)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_mor_all",
  start = "start",
  episode = "season"
)

```

### Full table
```{r echo=FALSE}
explanatory = c("vitd_supp*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat" ,"bst" ,"imd_bd_q", "regions", "low_immune", "chronic_disease")
explanatory_multi = c("vitd_supp*strata(season)", "sex", "age_group")

dependent ="Surv(start, end_mor_all, event = event)"

# model
bd_d_t4 %>%  finalfit(dependent, explanatory, explanatory_multi, keep_models = T) %>% knitr::kable(format = "pandoc", row.names = NA)
```

### Multiple regression
```{r eval=FALSE, include=FALSE}
coxph(Surv(start, end_mor_all, event = event) ~ vitd_supp*strata(season) + sex + age_group + ethnic_c + bmi_group + drink_freq_c + smoke_stat + bst+imd_bd_q + regions + low_immune + chronic_disease, 
      data = bd_d_t4 , id = f.eid) %>% summary()
```

### Simple regression
```{r echo=FALSE}
coxph(Surv(start, end_mor_all, event = event)~vitd_supp*strata(season), data = bd_d_t4 , id = f.eid) %>% summary()
```

### Partial adjusted for sex and age
```{r echo=FALSE}
partial_sen_o3 <- function(x){
  bd_d_t4 %>% 
  finalfit(dependent ="Surv(start, end_mor_all, event = event)", 
           explanatory = c(x,"sex", "age_group"), keep_models = T) %>% 
    knitr::kable(format = "pandoc")
}

var_list <- c("ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat","bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_o3)
```

# Supplementation table 9: Lab-confirmed Covid-19
```{r include=FALSE}
# data management 

bd_d$fu_time_lab <-(as.numeric(bd_d$end_d_lab) - as.numeric(start))
# Split the following time:

bd_d_lab <- survSplit(
  Surv(bd_d$fu_time_lab, event = bd_d$c_result)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_d_lab",
  start = "start",
  episode = "season"
)

```

### Simple regression
```{r echo=FALSE}

coxph(Surv(start, end_d_lab, event = event)~vitd_s*strata(season), data = bd_d_lab, id = f.eid) %>% summary()

```

### Partial adjustment: sex and age group
```{r echo=FALSE}

# function
partial_sen_lab <- function(x){
  bd_d_lab %>% 
  finalfit(dependent ="Surv(start, end_d_lab, event = event)", 
           explanatory = c(x,"sex", "age_group")) %>% knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat","bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, partial_sen_lab)
```

### Complete multiple regression
```{r}
# stratified analysis
explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")
dependent ="Surv(start, end_d_lab, event = event)"
bd_d_lab %>% finalfit(dependent, explanatory) %>% knitr::kable( format = "pandoc")

```

# Subgroup analysis among patients with Covid-19 diagnosis

## Supplementation table 10: (Hospitalization) subgroup analysis among patients with Covid-19 diagnosis

Data management: time splitting
```{r include=FALSE}
bd_d$fu_time_t3 <- (as.numeric(bd_d$end_hos_all)-as.numeric(start))

# Split the following time:

bd_d_t3 <- survSplit(
  Surv(bd_d$fu_time_t3, event = bd_d$hos_all)~.,
  data = bd_d,
  cut = time_cut, 
  end = "end_hos_all",
  start = "start",
  episode = "season"
)

```

### Simple regression for vit_d (can't use finalfit)
```{r echo=FALSE}
coxph(Surv(start, end_hos_all, event = event) ~ vitd_s*strata(season), data = (bd_d_t3 %>% filter(c_infection ==1)) , id = f.eid) %>% 
summary()
```

### Paritial adjustment for sex and age
```{r echo=FALSE}
# write a function for all
time_split_t3_partial <- function(x){
bd_d_t3 %>% filter(c_infection ==1) %>% 
    finalfit(
  dependent = "Surv(start, end_hos_all, event = event)", 
  explanatory = c(x, "sex", "age_group")) %>% 
  knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, time_split_t3_partial)


```

### Complete simple and multiple regression

```{r echo=FALSE}
bd_d_t3 %>% filter(c_infection ==1) %>% 
  finalfit(
  dependent = "Surv(start, end_hos_all, event = event)",
  explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat","bst", "imd_bd_q" , "regions","low_immune","chronic_disease")) %>%
  knitr::kable(format = "pandoc", row.names = NA)
```

## Supplementation table 11: (Mortality) Subgroup analysis among patients with Covid-19 diagnosis

Data management:

```{r include=FALSE}
bd_d$fu_time_t4 <-(as.numeric(bd_d$end_mor_all) - as.numeric(start))
# Split the following time:

bd_d_t4 <- survSplit(
  Surv(bd_d$fu_time_t4, event = bd_d$n_death_covid_all)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_mor_all",
  start = "start",
  episode = "season"
)

```

### Simple regression (without finalfit)
```{r echo=FALSE}
coxph( 
  Surv(start, end_mor_all, event = event) ~ vitd_s*strata(season), 
  data = (bd_d_t4 %>% filter(c_infection ==1)) , 
  id = f.eid) %>% summary()
```

### Partial adjustment for sex and age
```{r echo=FALSE}
# write a function for map
time_split_t3_partial <- function(x){
  bd_d_t4  %>% filter(c_infection ==1) %>%
    finalfit(
  dependent = "Surv(start, end_mor_all, event = event)", 
  explanatory = c(x, "sex", "age_group")) %>% 
  knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat","bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, time_split_t3_partial)
```

### Complete simple and multiple regression

```{r echo=FALSE}
bd_d_t4  %>% filter(c_infection ==1) %>%
  finalfit(
  dependent = "Surv(start, end_mor_all, event = event)",
  explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst", "imd_bd_q" , "regions","low_immune","chronic_disease")) %>%
  knitr::kable(format = "pandoc", row.names = NA)
```

