---
title: 'Table 3: Primary outcome: stratified Cox regression'
author: "Liang-Yu Lin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r include=FALSE}
# Load data and packages
# Datasets

# packages
library(markdown)
library(rmarkdown)
library(arsenal)
library(kableExtra)
library(survival)
library(survminer)
library(finalfit) 

# Limit significant digits to 2, remove scientific notation
options(digits = 2, scipen = 999)

```

# Time splitting: split follow-up time using `survSplit()`

Use `survSplit()` from survival package to split the following time.

Important time points: - Follow-up start date: 2020-03-15 - End of the British summertime: 2020-10-25 - End of following-up: end_d in the dataset

1.  Calculate the duration (i.e. the cut point) of British summer time (bst):

```{r echo=FALSE}
start <- as.Date("2020-03-15")
# spring_end <-  as.Date("2020-06-20")
# summer_end <-  as.Date("2020-09-22")
# autumn_end <-  as.Date("2020-12-21") 
bst <- as.Date("2020-10-25")

# calculate the days between two dates
# julian(spring_end, start) #97
# julian(summer_end, start) #191
# julian(autumn_end, start) #281
julian(bst, start) # 224

time_cut <- c(224) # try using BST  
```

2.  Split the following time. This will end up with two following time.

```{r include=FALSE}
bd_d_s <- survSplit(
  Surv(bd_d$fu_time, event = bd_d$c_infection)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_d",
  start = "start",
  episode = "season"
)
```

Three variables were created: "start", "event", and "season."

# Table 3: time splitting

**Table 3. The association between vitamin D status and Covid-19 infection**

### Simple regression of vit_d (can't use finalfit)

```{r echo=FALSE}
coxph(Surv(start, end_d, event = event) ~ vitd_s*strata(season), data = bd_d_s, id = f.eid) %>% 
summary()
```

### Partial adjustment for sex and age

```{r echo=FALSE}
# write a function for map
time_split_t2_partial <- function(x){
  bd_d_s %>% finalfit(
  dependent = "Surv(start, end_d, event = event)", 
  explanatory = c(x, "sex", "age_group")) %>% 
  knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, time_split_t2_partial)

```

### Complete simple and multiple regression

```{r echo=FALSE}
explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c", "smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")
dependent ="Surv(start, end_d, event = event)"
bd_d_s %>% finalfit(dependent, explanatory) %>% knitr::kable( format = "pandoc")
```

# Table 4: time splitting for hospitalization

Data management: time splitting

```{r include=FALSE}
bd_d$fu_time_t3 <- (as.numeric(bd_d$end_hos_all)-as.numeric(start))

# Split the following time:

bd_d_t3 <- survSplit(
  Surv(bd_d$fu_time_t3, event = bd_d$hos_all)~.,
  data = bd_d,
  cut = time_cut, 
  end = "end_hos_all",
  start = "start",
  episode = "season"
)

```

**Table 4. The association between vitamin d status and hospitalization due to covid-19**

### Simple regression for vit_d (can't use finalfit)

```{r echo=FALSE}
coxph(Surv(start, end_hos_all, event = event) ~ vitd_s*strata(season), data = bd_d_t3 , id = f.eid) %>% 
summary()

```

### Paritial adjustment for sex and age

```{r echo=FALSE}
# write a function for all
time_split_t3_partial <- function(x){
bd_d_t3 %>% finalfit(
  dependent = "Surv(start, end_hos_all, event = event)", 
  explanatory = c(x, "sex", "age_group")) %>% 
  knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, time_split_t3_partial)


```

### Complete simple and multiple regression

```{r echo=FALSE}
bd_d_t3 %>% finalfit(
  dependent = "Surv(start, end_hos_all, event = event)",
  explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst", "imd_bd_q" , "regions","low_immune","chronic_disease")) %>%
  knitr::kable(format = "pandoc", row.names = NA)
```

# Table 5: time splitting

Data management:

```{r include=FALSE}
bd_d$fu_time_t4 <-(as.numeric(bd_d$end_mor_all) - as.numeric(start))
# Split the following time:

bd_d_t4 <- survSplit(
  Surv(bd_d$fu_time_t4, event = bd_d$n_death_covid_all)~.,
  data = bd_d,
  cut = time_cut,
  end = "end_mor_all",
  start = "start",
  episode = "season"
)

```

**Table 5 : the association between vitamin D status and COVID-19 mortality**

### Simple regression (without finalfit)

```{r echo=FALSE}
coxph( 
  Surv(start, end_mor_all, event = event) ~ vitd_s*strata(season), 
  data = bd_d_t4 , 
  id = f.eid) %>% summary()
```

### Partial adjustment for sex and age

```{r echo=FALSE}
# write a function for map
time_split_t3_partial <- function(x){
  bd_d_t4 %>% finalfit(
  dependent = "Surv(start, end_mor_all, event = event)", 
  explanatory = c(x, "sex", "age_group")) %>% 
  knitr::kable( format = "pandoc")
}

var_list <- c("vitd_s*strata(season)","ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst","imd_bd_q" , "regions", "low_immune", "chronic_disease")

map(var_list, time_split_t3_partial)
```

### Complete simple and multiple regression

```{r echo=FALSE}
bd_d_t4 %>% finalfit(
  dependent = "Surv(start, end_mor_all, event = event)",
  explanatory = c("vitd_s*strata(season)", "sex", "age_group", "ethnic_c" , "bmi_group" , "drink_freq_c","smoke_stat", "bst", "imd_bd_q" , "regions","low_immune","chronic_disease")) %>%
  knitr::kable(format = "pandoc", row.names = NA)
```
