---
title: "Results table 2, summplementary table 1-2 and model checking"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
author: "Liang-Yu Lin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

**Load data and packages** 
```{r include=FALSE}
# Datasets
source('main_dataset.R')

# packages
library(markdown)
library(rmarkdown)
library(arsenal)
library(kableExtra)
library(finalfit) 
library(survival)
library(survminer)
library(epiDisplay)

# Limit significant digits to 2, remove scientific notation
options(digits = 2, scipen = 999)

```

# Table 1.
**Table 1. The distribution of demographic characteristics by vitamin D status**

```{r echo=FALSE}
t1vd <- arsenal::tableby( vitd_s ~ sex + age_group + ethnic_c + bmi_group + drink_freq_c + drink_freq_c + drink_stat_c + smoke_stat + imd_bd_q + vitd_supp + vd_drug + bst+ regions + low_immune + chronic_disease, 
                          data = bd_d,
                          cat.stats = "countpct"
                                )
summary(t1vd, text = T) %>% 
        kable(format = "pandoc",
              caption = "Formatted table",
              padding = 0,
              label = "",
              align = c("l","c"),
              col_order = order
        )

```

# Supplementary Table 1. 

Supplementary table 1. The comparison of inclusion and exclusion participants

```{r echo=FALSE}
# Read the original data
bd <- read_rds("V:/VOLUMEO/uk_biobank/1.work_data/ukb_study_2/datasets/complete_bd.rds")

# compare
include <- bd_d$f.eid %>% unlist()
bd %<>% mutate(exclude=ifelse(f.eid %in% include,0,1))
bd$exclude %<>% factor(labels = c("Included", "Excluded"))

# as.factor
bd_clinical_cov[,c("low_immune", "chronic_disease")] <- map(bd_clinical_cov[,c("low_immune", "chronic_disease")], as.factor)

bd %<>% left_join(bd_clinical_cov, by = c("f.eid" = "f.eid"))

bd %<>% mutate(vd_drug=ifelse(
  f.eid %in% unlist(bd_vd_drug$f.eid), 1 , 0))
bd$vd_drug<- bd$vd_drug %>% as.factor() %>% factor(labels = c("Not having vitamin D prescription","Had vitamin D prescription"))

ts1_control <- tableby.control(test = F, total = F, numeric.stats = c("N", "sd"), cat.stats=c("Nmiss","countpct")
)
# Change the percentage into column percentage
arsenal::tableby( exclude ~ sex + age_group + ethnic_c + bmi_group + drink_freq_c + drink_freq_c + drink_stat_c + smoke_stat + imd_bd_q + vitd_supp + vd_drug + regions + low_immune + chronic_disease, 
                          data = bd,
                          control = ts1_control
                                ) %>% 
summary(text = T) %>% 
        kable(format = "pandoc",
              caption = "Formatted table",
              padding = 0,
              label = "",
              align = c("l","c"),
              col_order = order
        )

```

# Testing proportional hazard assumption for the primary exposure

**Settings**
1.Define risk set / survival object
```{r include=FALSE}
riskset <- bd_d %$% Surv(fu_time, event = c_infection)
```
Cox regression by vitamin D status:
```{r include=FALSE}
# crude analysis
(cox_vd_s <- coxph(riskset ~ vitd_s, data = bd_d, id = f.eid))
confint(cox_vd_s,level = 0.95) %>% exp()
```

## Log-log(time) plot

### vitamin D status

```{r echo=FALSE, fig.height=3, fig.width=4}
# Use log log log(time) plot (crude)

log_o1 <- ggsurvplot(
  fit = survfit(riskset ~ bd_d$vitd_s), 
  data = bd_d, 
  fun = "cloglog", loglog =T,
  conf.int = F,
  xlim = c(1, 350),
  main = "Log-Log Plots",
  ggtheme = theme_bw(),
  legend = c("bottom"),
  legend.title = "Vitamin D status",
  legend.labs= c("Deficiency","Insufficiency","Sufficiency")
)

log_o1$plot <- log_o1$plot + 
  geom_vline(xintercept = 78, linetype = "dashed")+
  annotate("text", x = 78, y=-10, label = "Primary school open (2020-06-01)", hjust = "right", size = 6)+
  geom_vline(xintercept = 100, linetype = "dashed")+
  annotate("text", x = 100, y=-7.5, label = "The first lockdown ease",hjust = 0.2, size = 6)+
  annotate("text", x = 100, y=-8, label = "(2020-06-23)",hjust = 0.2, size = 6)+
  geom_vline(xintercept = 169, linetype = "dashed")+
  annotate("text", x = 169, y=-2.5, label = "Eat out help out ended", size = 6)+
  annotate("text", x = 169, y=-3, label = "(2020-08-31)", size = 6)+
  geom_vline(xintercept = 265, linetype = "dashed")+
  annotate("text", x = 265, y=0, label = "The second lockdown ease", size = 6) +
  annotate("text", x = 265, y=-0.5, label = "(2020-12-02)", size = 6)


ggpubr::ggpar(log_o1, font.legend = list(size = 14, color = "black", face = "bold"))

```

**Interpretation: The proportional hazard assumption is violated.**

### Hospitalization

```{r include=FALSE}
# survival objects: 
# time-in:  2020-03-15
# time-out: end_hos_all
bd_d$fu_time_t3 <- (as.numeric(bd_d$end_hos_all)-as.numeric(start))

riskset_hos <- bd_d  %$% Surv(fu_time_t3, event = hos_all)
```

```{r eval=FALSE, include=FALSE}
cox_hos_crude <- coxph(riskset_hos ~ vitd_s, data = bd_d, id = f.eid)
cox_hos_adj <- coxph(riskset_hos ~ vitd_s + sex + age_group +ethnic_c + bmi_group + drink_freq_c + imd_bd_q +  bst +regions + low_immune + chronic_disease, data =bd_d, id = f.eid)
cox.zph(cox_hos_crude)
cox.zph(cox_hos_adj)
```

Log log plot

```{r echo=FALSE, fig.height=3, fig.width=4}
# log log plot
log_o2 <- ggsurvplot(
  fit = survfit(riskset_hos ~ bd_d$vitd_s), 
  data = bd_d, 
  fun = "cloglog", loglog =T,
  conf.int = F,
  xlim = c(1, 350),
  main = "Log-Log Plots",
  ggtheme = theme_bw(),
  legend = c("bottom"),
  legend.title = "Vitamin D status",
  legend.labs= c("Deficiency","Insufficiency","Sufficiency")
)

log_o2$plot <- log_o2$plot + 
  geom_vline(xintercept = 78, linetype = "dashed")+
  annotate("text", x = 78, y=-10, label = "Primary school open (2020-06-01)", hjust = "right", size = 6)+
  geom_vline(xintercept = 100, linetype = "dashed")+
  annotate("text", x = 100, y=-7.5, label = "The first lockdown ease",hjust = 0.2, size = 6)+
  annotate("text", x = 100, y=-8, label = "(2020-06-23)",hjust = 0.2, size = 6)+
  geom_vline(xintercept = 169, linetype = "dashed")+
  annotate("text", x = 169, y=-2.5, label = "Eat out help out ended", size = 6)+
  annotate("text", x = 169, y=-3, label = "(2020-08-31)", size = 6)+
  geom_vline(xintercept = 265, linetype = "dashed")+
  annotate("text", x = 265, y=0, label = "The second lockdown ease", size = 6) +
  annotate("text", x = 265, y=-0.5, label = "(2020-12-02)", size = 6)


ggpubr::ggpar(log_o2, font.legend = list(size = 14, color = "black", face = "bold"))
```

**Interpretation: The proportional hazard assumption is violated.**

### Mortality

```{r include=FALSE}
# survival objects:
# time-in: 2020-03-15
# time-out: end_mor
bd_d$fu_time_t4 <-(as.numeric(bd_d$end_mor_all) - as.numeric(start))
riskset_mor <- bd_d %$% Surv(fu_time_t4, event =n_death_covid_all)

```

```{r eval=FALSE, include=FALSE}
cox_mor_crude <- coxph(riskset_mor ~ vitd_s, data = bd_d, id = f.eid)
cox_mor_adj <- coxph(riskset_mor ~ vitd_s + sex + age_group +ethnic_c + bmi_group + drink_freq_c + imd_bd_q +  bst +regions + low_immune + chronic_disease, data =bd_d, id = f.eid)
cox.zph(cox_mor_crude)
cox.zph(cox_mor_adj)
```

Log log plot

```{r echo=FALSE, fig.height=3, fig.width=4}
# log log plot
log_o3 <- ggsurvplot(
  fit = survfit(riskset_mor ~ bd_d$vitd_s), 
  data = bd_d, 
  fun = "cloglog", loglog =T,
  conf.int = F,
  xlim = c(1, 350),
  main = "Log-Log Plots of mortality",
  ggtheme = theme_bw(),
  legend = c("bottom"),
  legend.title = "Vitamin D status",
  legend.labs= c("Deficiency","Insufficiency","Sufficiency")
)

log_o3$plot <- log_o3$plot + 
  geom_vline(xintercept = 78, linetype = "dashed")+
  annotate("text", x = 78, y=-10, label = "Primary school open (2020-06-01)", hjust = "right", size = 6)+
  geom_vline(xintercept = 100, linetype = "dashed")+
  annotate("text", x = 100, y=-7.5, label = "The first lockdown ease",hjust = 0.2, size = 6)+
  annotate("text", x = 100, y=-8, label = "(2020-06-23)",hjust = 0.2, size = 6)+
  geom_vline(xintercept = 169, linetype = "dashed")+
  annotate("text", x = 169, y=-2.5, label = "Eat out help out ended", size = 6)+
  annotate("text", x = 169, y=-3, label = "(2020-08-31)", size = 6)+
  geom_vline(xintercept = 265, linetype = "dashed")+
  annotate("text", x = 265, y=0, label = "The second lockdown ease", size = 6) +
  annotate("text", x = 265, y=-0.5, label = "(2020-12-02)", size = 6)


ggpubr::ggpar(log_o3, font.legend = list(size = 14, color = "black", face = "bold"))
```

**Interpretation: The proportional hazard assumption is violated.**

# Supplementary table 2

Test the association between BST and vitamin D status
```{r echo=FALSE}
# Dataset:use the complete dataset

# 1. Data management: -----------------------------------------
# a. Exclude people without vitamin D test
bd_t <- bd %>% filter(vd_miss == 1)

# b. Recode vitamin D into vd25 and vd50
tab1(bd_t$vitd_s)

bd_t <- bd_t %>% 
        mutate(vd25=ifelse(vitd_s=="0_deficiency", 1, 0)) %>%  # vd25=deficiency
        mutate(vd50= ifelse(vitd_s=="0_deficiency" | vitd_s=="1_insufficiency", 1,0)) 
#vd50=insufficiency + deficiency


# Date issue: recode the BST date 

# Roughly recode the BST
bd_t %<>% mutate(bst=case_when(
        month(date_0) %in% c(4,5,6,7,8,9,10) ~ 1,
        month(date_1) %in% c(4,5,6,7,8,9,10) ~ 1,
        month(date_0) %in% c(11,12,1,2,3) ~ 0,
        month(date_1) %in% c(11,12,1,2,3) ~ 0,
))

bd_t$bst %<>% as.factor()
bd_t$bst %<>% factor(labels = c("Not_in_bst_months", "During_bst_months"))
bd_t$bst %>% tab1
# Not_in_bst_months    169058    37.6         37.6
# During_bst_months    280880    62.4        100.0
# Total              449938   100.0        100.0


# 2. Statistical analysis ------

# Crude logistic regression
# vd25:
# crude:
vd25_crude <- glm(vd25 ~ bst, data = bd_t, family = "binomial")
vd25_crude %>% logistic.display()
# adjusted for covariates
vd25_adj <- glm(vd25 ~ bst + age_c + regions + ethnic + bmi_group + 
                        imd_bd_q + drink_freq_c + smoke_stat, 
                data = bd_t, family = "binomial")
vd25_adj %>% logistic.display() 


# vd50
# crude:
vd50_crude <- glm(vd50 ~ bst, data = bd_t, family = "binomial")
vd50_crude %>% logistic.display()

# adjusted for covariates
vd50_adj <- glm(vd50 ~ bst + age_c + regions + ethnic + bmi_group + 
                        imd_bd_q + drink_freq_c + smoke_stat,
                data = bd_t, family = "binomial")
vd50_adj %>% logistic.display()

```

